Analysis for Case 2 (habituals)

### 2018-02-08 analysis/case2-habituals.Rmd
port over more code into scripts

# gen.inset.distributions <- bind_rows(
#   gen.listener.predictions %>%
#     mutate(category = NA),
#   m.gen.speakerBeliefs %>%
#     select(property, val) %>%
#     rename(value = val) %>%
#     mutate(Parameter = "speakerBeliefs")
# )
# 
# 
# category.text.labels <- data.frame(property = c("dont eat people", 
#                       "carry malaria", "lay eggs",  "are female", "have spots"),
#              category = c("Tigers", "Mosquitos", "Robins", "Robins", "Leopards"),
#              x = c(0.3, 0.3, 0.47, 0.05, 0.6),
#             y = c(0.45, 0.5, 0.45, 0.5, 0.26))
```

```{r}
save(m.hab.endorse.rsa, hab.uncertain.threshold.s1opt, fixed.hab.params.posterior, md.hab.long, hab.listener.predictions,
     file = "../paper/cached_results/case2_endorsement_rsaModelResults.RData")

@@

>>>>

# gen.inset.distributions <- bind_rows(
#   gen.listener.predictions %>%
#     mutate(category = NA),
#   m.gen.speakerBeliefs %>%
#     select(property, val) %>%
#     rename(value = val) %>%
#     mutate(Parameter = "speakerBeliefs")
# )
# 
# 
# category.text.labels <- data.frame(property = c("dont eat people", 
#                       "carry malaria", "lay eggs",  "are female", "have spots"),
#              category = c("Tigers", "Mosquitos", "Robins", "Robins", "Leopards"),
#              x = c(0.3, 0.3, 0.47, 0.05, 0.6),
#             y = c(0.45, 0.5, 0.45, 0.5, 0.26))
hab.listener.predictions

m.hab.endorse.rsa
hab.uncertain.threshold.s1opt
fixed.hab.params.posterior
md.hab
md.hab.long
md.hab
hab.listener.predictions
```

```{r saveModels}
save(m.hab.endorse.rsa, hab.uncertain.threshold.s1opt, fixed.hab.params.posterior,
     md.hab.long, md.hab, hab.listener.predictions,
     file = "../paper/cached_results/case2_endorsement_rsaModelResults.RData")
```

@@

## Experiment 2c: What is prevalence?

```{r habituals-predictive-elicitation}
d.hab.predictive <- read.csv(paste(
  project.path, "data/habituals/predictive/", "predictive-1-trials.csv", sep = ""))

n.subj.hab.pred <- length(unique(d.hab.predictive$workerid))
ave.seconds.hab.pred <-  d.hab.predictive %>% select(workerid, rt) %>%
      group_by(workerid) %>% summarize(totalSec= sum(rt) / 1000) %>% ungroup() %>% 
      summarize(total = mean(totalSec))

# participant reported this in the comments section (it was recorded as a symbol)
d.hab.predictive[
  (d.hab.predictive$workerid==44 & 
     d.hab.predictive$item=="smokes cigarettes"),"response"] <- 5
d.hab.predictive <- d.hab.predictive %>% mutate(response = as.numeric(as.character((response))))

d.hab.predictive<- d.hab.predictive %>%
  rowwise() %>%
  mutate(
    annualPredictiveRate =
           annualRates[[as.character(past_interval)]]*response,
    annualPastRate =
           annualRates[[as.character(past_interval)]]*past_freq,
    annualPredictiveRate = ifelse(
      annualPredictiveRate == 0, 0.05, 
      annualPredictiveRate),
    logAnnualPredictiveRate = log(annualPredictiveRate),
    logAnnualPastRate = log(annualPastRate)
    )

d.hab.predictive.summary <- d.hab.predictive %>%
  group_by(condition, item, logAnnualPastRate) %>%
  tidyboot_mean(column = logAnnualPredictiveRate)

r.hab.pred.past.baseline.freq <- round(with(d.hab.predictive.summary %>% filter(condition == "baseline"), cor(logAnnualPastRate, mean)), 3)
n.hab.pred.past.baseline.freq <- length(filter(d.hab.predictive.summary, condition == "baseline")$mean)

fig.hab.freq.predictive.vs.past <- ggplot(d.hab.predictive.summary,
       aes(x=logAnnualPastRate,
                              y = mean,
                              ymin = ci_lower, ymax = ci_upper,
                              fill = condition, shape=condition))+
  geom_jitter(position = position_jitter(width = .13), alpha = 0.6, size = 4)+
  scale_shape_manual(values=c(21,22,23))+
  geom_abline(intercept = 0, slope = 1, lty =3, color = 'black')+
  # #geom_errorbar(width=0.1)+
  scale_x_continuous(limits = c(-3,6),
                     breaks = c(-3, 0, 2.5, 4, 5.9),
                     labels = c("almost never", "annually", "monthly", "weekly", "daily"))+
  scale_y_continuous(limits = c(-3,6),
                     breaks = c(-3, 0, 2.5, 4, 5.9),
                     labels = c("almost never", "annually", "monthly", "weekly", "daily"))+  coord_fixed()+
  guides(fill=F, shape=F)+
  scale_fill_solarized()+
  xlab("Past frequency (log scale)")+
  ylab("Predicted frequency (log scale)")+
  theme(axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1))

d.hab.predictive.summary
```

```{r habituals-predictive-model}
hab.predictive.bda.model <- '
var model = function(){

	var predictiveFrequency = {
			mu: uniformDrift({a: -5, b:10, width: 2}),
			sigma: uniformDrift({a:0, b:10, width: 1})
	}

	mapData({data: data}, function(d){
		observe(Gaussian(predictiveFrequency),
        d.logAnnualPredictiveRate)
	})

  return predictiveFrequency
}
'
```

```{r habituals-predictive-model-run, cache = T, eval=F}
items <- levels(factor(d.hab.predictive$item))
conditions <- levels(factor(d.hab.predictive$condition))
n_samples <- 5000
rs.habituals.bda.predictive <- data.frame()

for (it in items){
  for (cn in conditions){
     df.predictive.toPass <- d.hab.predictive %>%
       filter(condition == cn, item == it)

    rs <- webppl(program_code = hab.predictive.bda.model,
       model_var = "model",
       inference_opts = list(method = "MCMC", samples = n_samples,
                             burn = n_samples / 2),
       data = df.predictive.toPass,
       data_var = "data")
    rs.habituals.bda.predictive <- bind_rows(
        rs.habituals.bda.predictive, 
        rs %>% 
            mutate(item = it, condition = cn) 
        )
  }
}
```

```{r eval = F}
  d.hab.predictive %>% 
    mutate(src = 'data') %>% 
    select(item, condition, src, logAnnualPredictiveRate) %>%
ggplot(. , aes( x = logAnnualPredictiveRate, fill = condition))+
#  geom_density(size = 1, aes( y = ..scaled.. ))+
  geom_histogram(aes(y=(..count..)/tapply(..count..,..PANEL..,sum)[..PANEL..]),
               position = position_dodge())+
  facet_wrap(~item, scales = 'free')+
  scale_color_solarized()+
  #scale_x_continuous(limits = c(-0.05,1.05), breaks = c(0, 0.5, 1)) +
  scale_x_continuous(limits = c(-5,10)) +
  theme(strip.text.y = element_text(angle = 0))
```

```{r habituals-predictive-bda-forwardSample, fig.width=14, eval=F}
# rs.habituals.bda.predictive.samples <- rs.habituals.bda.predictive %>% 
#   spread(Parameter, value) %>%
#   rowwise() %>%
#   mutate(
#     logannualRate = rnorm(n = 1, mean = mu, sd = sigma)
#   )

  d.hab.predictive %>% 
    mutate(src = 'data') %>% 
    select(item, condition, src, logAnnualPredictiveRate) %>%
ggplot(. , aes( x = logAnnualPredictiveRate, color = condition))+
#  geom_density(size = 1, aes( y = ..scaled.. ))+
  geom_histogram(aes(y=(..count..)/tapply(..count..,..PANEL..,sum)[..PANEL..]),
               position = position_dodge())+
  facet_wrap(~item, scales = 'free')+
  scale_color_solarized()+
  #scale_x_continuous(limits = c(-0.05,1.05), breaks = c(0, 0.5, 1)) +
  scale_x_continuous(limits = c(-5,10)) +
  theme(strip.text.y = element_text(angle = 0))

bind_rows(
  d.hab.predictive %>% 
    mutate(src = 'data') %>% 
    select(item, condition, src, logAnnualPredictiveRate),
  rs.habituals.bda.predictive.samples %>% 
    select(item, condition, logannualRate) %>%
    rename(logAnnualPredictiveRate = logannualRate) %>%
    mutate(src = 'model') 
  ) %>%
ggplot(. , aes( x = logAnnualPredictiveRate, color = src))+
  geom_density(size = 1, aes( y = ..scaled.. ))+
#  geom_histogram(aes(y=(..count..)/tapply(..count..,..PANEL..,sum)[..PANEL..]),
#               position = position_dodge())+
  facet_grid(condition ~ item )+
  scale_color_solarized()+
  scale_x_continuous(limits = c(-5,10)) +
  #scale_y_continuous(limits = c(-0.01,1.01), breaks = c(0, 0.5, 1)) +
  theme(strip.text.y = element_text(angle = 0))
```

```{r habituals-predictive-bda-summary, eval =F}
rs.habituals.bda.predictive.summary <- rs.habituals.bda.predictive %>% 
  spread(Parameter, value) %>%
  rowwise() %>%
  mutate(
    logannualRate = rnorm(n = 1, mean = mu, sd = sigma)
  ) %>%
  gather(key, val, mu, sigma, logannualRate) %>%
  group_by(item, condition, key) %>%
  summarize(MAP = estimate_mode(val),
            cred_upper = hdi_upper(val),
            cred_lower = hdi_lower(val))
```

```{r habituals-predictive-bda-figure, eval =F}
fig.hab.freq.predictive.vs.past <- left_join(
  rs.habituals.bda.predictive.summary %>% filter(key == "logannualRate"),
  unique(d.hab.predictive %>% 
    select(item, condition, logAnnualPastRate)
  )) %>% ggplot(., 
       aes(x=logAnnualPastRate, y = MAP, 
                              ymin = cred_lower, ymax = cred_upper, 
                              fill = condition, shape=condition))+
  geom_jitter(position = position_jitter(width = .13), alpha = 0.6)+
  scale_shape_manual(values=c(21,22,23, size = 4))+
  geom_abline(intercept = 0, slope = 1, lty =3, color = 'black')+
  # #geom_errorbar(width=0.1)+
  scale_x_continuous(limits = c(-4,7), 
                     breaks = c(-4, 0, 2.5, 4, 5.9),
                     labels = c("almost never", "annually", "monthly", "weekly", "daily"))+
  scale_y_continuous(limits = c(-4,7), 
                     breaks = c(-4, 0, 2.5, 4, 5.9),
                     labels = c("almost never", "annually", "monthly", "weekly", "daily"))+  coord_fixed()+
  guides(fill=F, shape=F)+
  scale_fill_solarized()+
  xlab("Past frequency (log scale)")+
  ylab("Predicted frequency MAP (log scale)")+
  theme(axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1))
```

```{r habituals-predictive-endorsement}
d.hab.predictive.endorsment <- read.csv(paste(
  project.path, "data/habituals/endorsement/", "habituals-endorsement-predictives.csv", sep = ""))
d.hab.predictive.endorsment.catch <- read.csv(paste(
  project.path, "data/habituals/endorsement/", "habituals-endorsement-predictives-catch_trials.csv", sep = ""))

n.subj.hab.pred.endorse <- length(unique(d.hab.predictive.endorsment$workerid))
ave.seconds.hab.pred.endorse <-  d.hab.predictive.endorsment %>% select(workerid, rt) %>%
      group_by(workerid) %>% summarize(totalSec= sum(rt) / 1000) %>% ungroup() %>% 
      summarize(total = mean(totalSec))

d.hab.predictive.endorsment.condition.summary <- left_join(
  d.hab.predictive.endorsment, 
  d.hab.predictive.endorsment.catch %>% select(workerid, pass),
  by = "workerid"
  ) %>% filter(pass==1) %>%
  mutate(response = ifelse(response=="agree-key", 1, 0)) %>%
  rename(item = habitual) %>%
  group_by(condition) %>%
  summarize(k = sum(response), n = n()) %>%
  ungroup() %>%
  mutate(a = 1 + k,
         b = 1 + n - k,
         low  = qbeta(.025, a, b),
         high = qbeta(.975, a, b),
         MAP_h = (a-1)/(a+b-2),
         mean = a / (a + b))

d.hab.predictive.endorsment.baseline.summary <- d.hab.predictive.endorsment.condition.summary %>%
  filter(condition == 'baseline')

d.hab.predictive.endorsment.preventative.summary <- d.hab.predictive.endorsment.condition.summary %>%
  filter(condition == 'preventative')

d.hab.predictive.endorsment.enabling.summary <- d.hab.predictive.endorsment.condition.summary %>%
  filter(condition == 'enabling')

d.hab.predictive.endorsment.summary <- left_join(
  d.hab.predictive.endorsment, 
  d.hab.predictive.endorsment.catch %>% select(workerid, pass),
  by = "workerid"
  ) %>% filter(pass==1) %>%
  mutate(response = ifelse(response=="agree-key", 1, 0),
         time_period = factor(time_period,
                              levels=c("month",
                                       "2 months",
                                       "year",
                                       "2 years",
                                       "5 years")),
         condition = factor(condition,
                            levels=c("enabling", "baseline", "preventative")),
         hab_time = paste(habitual, time_period, sep='-')) %>%
  rename(item = habitual) %>%
  group_by(item, condition, hab_time, time_period) %>%
  summarize(k = sum(response), n = n()) %>%
  ungroup() %>%
  mutate(a = 1 + k,
         b = 1 + n - k,
         low  = qbeta(.025, a, b),
         high = qbeta(.975, a, b),
         MAP_h = (a-1)/(a+b-2),
         mean = a / (a + b))

d.hab.predictive.endorsment.summary <- left_join(
  d.hab.predictive.endorsment.summary,
  d.hab.predictive.summary %>%
    rename(logAnnualFutureRate = mean,
         future_lower = ci_lower, future_upper = ci_upper) %>%
    select(-n)
  # rs.habituals.bda.predictive.summary %>% 
  #   filter(key == "logannualRate") %>%
  #   rename(logAnnualFutureRate = MAP,
  #        future_lower = cred_lower, future_upper = cred_upper)
  # %>%
  #   rowwise() %>%
  #   mutate(future_lower = ifelse(future_lower < -4, -4, future_lower),
  #          future_upper = ifelse(future_upper > 7, 7, future_upper))
)

r2.habituals.predictiveFreq <- compute_r2(d.hab.predictive.endorsment.summary,
                                          "logAnnualFutureRate", "MAP_h")
mse.habituals.predictiveFreq <- compute_mse(d.hab.predictive.endorsment.summary,
                                          "logAnnualFutureRate", "MAP_h")

fig.hab.endorse.vs.predfreq <- d.hab.predictive.endorsment.summary %>%
  ggplot(., aes(x=logAnnualFutureRate, y=MAP_h, 
                 fill=condition, shape=condition,
                 xmin = future_lower, xmax=future_upper,
                 ymin=low, ymax = high))+
  geom_errorbar(alpha=0.8)+
  geom_errorbarh(alpha=0.8)+
  scale_shape_manual(values=c(21,22,23,24))+
  geom_point(alpha = 1, size = 4)+
  xlab("Predicted frequency (log scale)")+
  ylab("Human endorsement")+
  scale_x_continuous(limits = c(-4.2,7.2), 
                     breaks = c(-3, 0, 2.5, 4, 5.9),
                     labels = c("almost never", "annually", "monthly", "weekly", "daily"))+
  scale_y_continuous(limits = c(0, 1), breaks = c(0, 0.5, 1))+
  coord_fixed(ratio=11.4)+
  scale_fill_solarized()+
  guides(fill = F, shape = F)+
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1),
    legend.position="bottom",
    legend.direction="horizontal"
    )

```{r habituals-predictive-rsa}
n_chains <- 3
# n_samples <- 500000
# burn <- n_samples / 2
# lg <- 150
model_prefix <- "results-habituals-predictive-jointModel-S1-smtncs_habitual-"
model_prefix <- "results-habituals-predictive-meanPredictive-min-5-jointModel-S1-smtncs_habitual-"
n_samples <- 100000
burn <- n_samples / 2
lg <- 50

m.hab.pred.samp <- data.frame()
m.hab.past.samp <- data.frame()

# for (i in seq(1, n_chains)){
#   mi <- fread(paste(project.path,  "models/habituals/results/",
#                     model_prefix, "targetH_predictive-",
#                     n_samples, "_burn", burn, "_lag", lg, "_chain", i, ".csv", sep = ""))
#   m.hab.pred.samp <- bind_rows(m.hab.pred.samp, mi %>% mutate(chain = i))

  # mi.past <- fread(paste(project.path,  "models/habituals/results/",
  #                   model_prefix, "targetH_past-",
  #                   n_samples, "_burn", burn, "_lag", lg, "_chain", i, ".csv", sep = ""))
  # m.hab.past.samp <- bind_rows(m.hab.past.samp, mi.past %>% mutate(chain = i))

#}
# 
# 
# # 
# save(m.hab.pred.samp,
#      file = paste(project.path,  "models/habituals/results/", model_prefix, "predictive-",
#                     n_samples, "_burn", burn, "_lag", lg, "_",n_chains , "chains.RData", sep = ""))
# # # 
# save(m.hab.past.samp,
#      file = paste(project.path,  "models/habituals/results/", model_prefix, "past-",
#                     n_samples, "_burn", burn, "_lag", lg, "_",n_chains , "chains.RData", sep = ""))

load(file = paste(project.path,  "models/habituals/results/results-habituals-predictive-jointModel-S1-smtncs_habitual-predictive-100000_burn50000_lag50_3chains.RData", sep = ""))

load(file = paste(project.path,  "models/habituals/results/results-habituals-predictive-jointModel-S1-smtncs_habitual-past-100000_burn50000_lag20_3chains.RData", sep = ""))

m.hab.pastmodel.endorsement <- m.hab.past.samp %>%
  filter(type == 'predictive', C == "endorsement") %>%
  rename(item = B, condition = D, binned_freq = E) %>%
  group_by(item, condition) %>%
  summarize(MAP = estimate_mode(val),
            cred_upper = hdi_upper(val),
            cred_lower = hdi_lower(val))

m.hab.predictive.endorsement <- m.hab.pred.samp %>%
  filter(type == 'predictive', C == "endorsement") %>%
  rename(item = B, condition = D, binned_freq = E) %>%
  group_by(item, condition) %>%
  summarize(MAP = estimate_mode(val),
            cred_upper = hdi_upper(val),
            cred_lower = hdi_lower(val))

m.hab.predictive.endorse.rsa <- bind_rows(
  left_join(
    d.hab.predictive.endorsment.summary,
    m.hab.predictive.endorsement
  ) %>%
    mutate(src = "predictive_model"),
  left_join(
    d.hab.predictive.endorsment.summary,
    m.hab.pastmodel.endorsement
  ) %>%
    mutate(src = "past_model")
)

r2.habituals.rsa.past <- compute_r2(
  m.hab.predictive.endorse.rsa %>% filter(src == "past_model"),
                                          "MAP_h", "MAP")

mse.habituals.rsa.past <- compute_mse(
  m.hab.predictive.endorse.rsa %>% filter(src == "past_model"),
                                          "MAP_h", "MAP")

r2.habituals.rsa.predictive <- compute_r2(
  m.hab.predictive.endorse.rsa %>% filter(src == "predictive_model"),
                                          "MAP_h", "MAP")

mse.habituals.rsa.predictive <- compute_mse(
  m.hab.predictive.endorse.rsa %>% filter(src == "predictive_model"),
                                          "MAP_h", "MAP")

r2.hab.rsa.predictive.n <- length(d.hab.predictive.endorsment.summary$MAP_h)

habituals.predictive.past.model <- m.hab.predictive.endorse.rsa %>%
  mutate(src = factor(src, levels = c( 
                                      "past_model",
                                      "predictive_model"
                                     ),
                      labels = c(
                                 "Past-frequency endorsement model",
                                 "Predictive-frequency endorsement model"

                                 ))) %>%
  filter(src == "Past-frequency endorsement model") %>%
  ggplot(., aes ( x = MAP, xmin = cred_lower, xmax = cred_upper,
                  y = MAP_h, ymin = low, ymax = high, fill = condition,
                  shape = condition))+
  geom_abline(intercept = 0, slope = 1, lty = 3)+
  geom_linerange(alpha = 0.7)+
  geom_errorbarh(alpha = 0.7)+
  geom_point(size = 4)+
  scale_shape_manual(values=c(21,22,23,24))+
  scale_x_continuous(limits = c(-0.01, 1.01), breaks = c(0,  1))+
  scale_y_continuous(limits = c(-0.01, 1.01), breaks = c(0, 1))+
  #scale_fill_continuous(low = "#2b83ba", high = "#d7191c")+
  coord_fixed()+
  xlab("Model prediction")+
  ylab("Human habitual endorsement")+
  facet_wrap(~src, nrow = 1)+
  guides(fill = F)+
  #theme(legend.position = "bottom")+
  scale_fill_solarized()

habituals.predictive.predictive.model <- m.hab.predictive.endorse.rsa %>%
  mutate(src = factor(src, levels = c( 
                                      "past_model",
                                      "predictive_model"
                                     ),
                      labels = c(
                                 "Past-frequency endorsement model",
                                 "Predictive-frequency endorsement model"

                                 ))) %>%
  #filter(src == "Predictive-frequency endorsement model") %>%
  ggplot(., aes ( x = MAP, xmin = cred_lower, xmax = cred_upper,
                  y = MAP_h, ymin = low, ymax = high, fill = condition, shape = condition))+
  geom_abline(intercept = 0, slope = 1, lty = 3)+
  geom_linerange(alpha = 0.7)+
  geom_errorbarh(alpha = 0.7)+
  geom_point(size = 4)+
  scale_shape_manual(values=c(21,22,23,24))+
  scale_x_continuous(limits = c(-0.01, 1.01), breaks = c(0,  1))+
  scale_y_continuous(limits = c(-0.01, 1.01), breaks = c(0, 1))+
  #scale_fill_continuous(low = "#2b83ba", high = "#d7191c")+
  coord_fixed()+
  xlab("\n Model prediction")+
  ylab("Human habitual endorsement")+
  facet_wrap(~src, nrow = 1)+
  #guides(fill = F)+
  #theme(legend.position = "bottom")+
  scale_fill_solarized()
#habituals.predictive.predictive.model
```

```{r fig.height = 6, fig.width=14, echo = F, eval=F}
m.hab.pred.samp %>%
  filter(type == "predictive", C == "predictiveFreq", E == "sigma") %>%
  #filter(type == "predictive", C == "endorsement") %>%
#  mutate(E = as.numeric(as.character(E))) %>%
  ggplot(., aes(x = val, fill = D))+
  geom_histogram(position = position_dodge())+
  #facet_grid(B~E)+
  facet_grid(D~B)
```

```{r, eval = F}
m.hab.predfreq.summary <- m.hab.pred.samp %>%
  filter(type == "predictive", C == "predictiveFreq") %>%
  group_by(B, D, E) %>%
  summarize(MAP = estimate_mode(val)) %>% spread(E, MAP)

left_join(m.hab.predictive.endorse.rsa %>%
  mutate(src = factor(src, levels = c( 
                                      "past_model",
                                      "predictive_model"
                                     ),
                      labels = c(
                                 "Endorsement model (past frequency)",
                                 "Endorsement model (predictive frequency)"

                                 ))) %>%
  filter(src == "Endorsement model (predictive frequency)"), m.hab.predfreq.summary %>% rename(item = B, condition = D, freqmean = mean)) %>% View()
```

```{r habituals-predictive-elicitation-predictions}
lmer.rs.hab.predictive <- lmer(data = d.hab.predictive, 
     logAnnualPredictiveRate ~ logAnnualPastRate + condition + 
       (1 + condition | workerid) + 
       (1 + condition | item)
)
lmer.rs.hab.predictive.summary <- summary(lmer.rs.hab.predictive)
```

```{r}
save(d.hab.predictive.endorsment.summary, d.hab.predictive.summary, 
     n.subj.hab.pred.endorse, ave.seconds.hab.pred.endorse, 
     n.subj.hab.pred, ave.seconds.hab.pred, d.hab.predictive.endorsment.condition.summary, 
     lmer.rs.hab.predictive.summary,
     file = "../paper/cached_results/case2_predictive_data.RData")
```
